---
title: "CIA Factbook Mashup"
author: "Robert Farrow, John Weng, Zhu Mei, Haichen Liu, Huanzhe Wang"
date: "December 9, 2015"
output: html_document
---

## Load packages needed for this project

```{r}
library(XML)
library(RColorBrewer)
library(maps)
# R looks for a binary version of the package, which is not available. Instead we have to use the source code for installing this package
install.packages("RKML", repos = "http://www.omegahat.org/R", type = "source") 
library(RKML)
library(RCurl)
library(plyr)
library(sp)
library(maptools)
library(rgdal)
```

## Google Earth will need to be downloaded to view the final product
# http://www.google.com/earth/download/ge/agree.html

# Download the CIA Factbook
# http://jmatchparser.sourceforge.net/factbook/data/factbook.xml.gz

```{r}
factbookDOC = xmlParse("C:/Users/Robert/Desktop/STAT6306_McGee/Project/factbook.xml")
factbookRoot = xmlRoot(factbookDOC)
```

# Inspect the Factbook

```{r}
xmlName(factbookRoot)
xmlSize(factbookRoot)
```

## Create the popDF and infMortDF Dataframes
# We made several educated guesses to get to the data in the Tree Node

```{r}
table(names(factbookRoot))
sapply(factbookRoot["category"], function(node) table(names(node)))
sapply(factbookRoot["category"], xmlAttrs)
categoryNodes = factbookRoot["category"]
w = sapply(categoryNodes, xmlGetAttr, "name")=="People and Society"
Ids = sapply(categoryNodes[[ which(w) ]] [ "field" ], xmlGetAttr, "id")
f2091Index = which(Ids == "f2091")
f2091Index
```

# Create the infMortDF Dataframe

```{r}
rankNodes = categoryNodes[[ which(w) ]][ "field" ][[ f2091Index ]]["rank"]
xmlSize(rankNodes) # CIA Factbook direction has 223, we have 224
infMortNum = sapply(rankNodes, xmlGetAttr, "number")
infMortCtry = sapply(rankNodes, xmlGetAttr, "country")
head(infMortNum)
head(infMortCtry)
field2091 = getNodeSet(factbookDOC, "//field[@id='f2091']") # CIA Factbook directions has "factbookDoc", should be "factbookDOC"
xmlAttrs(field2091[[1]])
rankNodes = getNodeSet(factbookDOC, "//field[@id='f2091']/rank")
xmlAttrs(rankNodes[[1]])
infNum = as.numeric(sapply(rankNodes, xmlGetAttr, "number"))
infCtry = sapply(rankNodes, xmlGetAttr, "country")
infMortDF = data.frame(infMort = infNum, ctry = infCtry, stringsAsFactors = FALSE)
```

# Create the popDF Dataframe

```{r}
rankNodes = getNodeSet(factbookRoot, "//field[@id='f2119']/rank")
popNum = as.numeric(sapply(rankNodes, xmlGetAttr, "number"))
popCtry = sapply(rankNodes, xmlGetAttr, "country")
popDF = data.frame(pop = popNum, ctry = popCtry, stringsAsFactors = FALSE)
```

## Get the Lattitude and Longitude data frame
# Website to download the data
# http://dev.maxmind.com/static/csv/codes/country_latlon.csv

```{r}
latlonDF <- read.csv("C:/Users/Robert/Desktop/STAT6306_McGee/Project/country_latlon.csv", sep=",")
```

## Get the ISO 3166 codes
# https://github.com/obikag/iso-3166-country-codes

```{r}
ISOCodes <- read.csv("C:/Users/Robert/Desktop/STAT6306_McGee/Project/iso&cia_country_codes.csv", sep=",")
```

## Combine the latlonDF, popDF, and infMortDF data frames

```{r}
head(latlonDF)
head(popDF)
head(infMortDF)
nrow(latlonDF)
IMPop = merge(infMortDF, popDF, by="ctry", all=FALSE)
dim(IMPop)
latlonDF$code=tolower(as.character(latlonDF$iso.3166.country))
latlonDFMerge=merge(latlonDF, ISOCodes, by="iso.3166.country", all=FALSE)
latlonDFMerge$ctry=tolower(as.character(latlonDFMerge$ctry))
allCtryData=merge(IMPop, latlonDFMerge, by.x="ctry", all=FALSE)
```

## Countries that did not contain values for either Population or Infant Moratality Rate

```{r}
IMPop1 = merge(infMortDF, popDF, by="ctry", all=TRUE)
is.na(IMPop1)
IMPop1[16,] #Ctry code is ax, Akrotiri
IMPop1[38,] #Ctry code is cc
IMPop1[46,] #Ctry code is ck, Cocos (Keeling) Islands
IMPop1[61,] #Ctry code is dx, Dhekelia
IMPop1[74,] #Ctry code is fk, Falkland Islands (Islas Malvinas)
IMPop1[117,] #Ctry code is kt, Christmas Island
IMPop1[119,] #Ctry code is kv, Kosovo
IMPop1[137,] #Ctry code is mj, Montenegro
IMPop1[151,] #Ctry code is ne, Niue
IMPop1[152,] #Ctry code is nf, Norfolk Island
IMPop1[165,] #Ctry code is pc, Pitcairn Islands
IMPop1[177,] #Ctry code is rn, Saint Martin
IMPop1[199,] #Ctry code is sv, Svalbard
IMPop1[203,] #Ctry code is tb, Saint Barthelemy
IMPop1[208,] #Ctry code is tl, Tokelau
IMPop1[231,] #Ctry code is vt, Holy See (Vatican City)
```

## Set the color palette and create the breaks to use for Infant Mortality Rate

```{r}
display.brewer.all()
# view the color palette
cols = brewer.pal(9, "YlOrRd")[c(1,2,4,6,7)]
newInfMort = cut(allCtryData$infMort, breaks = 5)
summary(newInfMort)
newInfMort2 = cut(allCtryData$infMort, breaks = c(0, 37, 50, 65, 80, 150))
summary(newInfMort2)
hist(allCtryData$infMort, breaks=20, main="", xlab="Infant Mortality per 1000 Live Births")
quantile(allCtryData$infMort, probs=seq(0, 1, by=.2))
InfMortDiscrete = cut(allCtryData$infMort, breaks=c(0, 10, 25, 50, 75, 150))
```

## Create popDiscrete

```{r}
quantile(allCtryData$pop, probs=seq(0, 1, by=.2))
popDiscrete = cut(allCtryData$pop, breaks=c(0, 320041, 3588317, 10127792, 29387171, 1355692577))
```

## Create the Map

```{r}
world = map(database="world", fill=TRUE, col="light grey")
symbols(allCtryData$longitude, allCtryData$latitude, add=TRUE, circles=sqrt(allCtryData$pop)/4000, inches=FALSE, fg=cols[InfMortDiscrete], bg=cols[InfMortDiscrete])
legend(x=-150, y=0, title="Infant Mortality", legend=levels(InfMortDiscrete), fill=cols, cex=.8)
```

## Check the range of the population and set the minimum radius

```{r}
range(allCtryData$pop)
hist(sqrt(allCtryData$pop), breaks=20, xlab="Square-root of Population", main="")
rads=pmax(sqrt(allCtryData$pop)/4000, 1)
```

## Plotting with Google Earth, the code in the text does not work

```{r}
doc=kmlPoints(allCtryData)
## The saveXML code crashes R, DO NOT USE!! Code provided by the text
# saveXML(doc, file = "C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/countryData1.kml")
```

## Our solution to the code in the text
## Create a Dataframe

```{r}
coordinates(allCtryData) <- ~latitude+longitude
proj4string(allCtryData) <- CRS("+proj=longlat +datum=WGS84")
```

## Create a Matrix of the coordinates

```{r}
rown<-nrow(coordinates(allCtryData))
coln<-ncol(coordinates(allCtryData))
Temp <- matrix(nrow=rown,ncol = 1)
id=1
for (i in 1:rown){
 Temp[i,1] <- id
  id<-id+1 
}
Temp
```

## Create the KML Document

```{r}
allCtryData.kml <- newXMLNode("kml")
h2 <- newXMLNode("Document", parent = allCtryData.kml)
h3 <- newXMLNode("name", "allCtryData", parent = h2)
h4 <- newXMLNode("LookAt", parent=h2)
h5 <- newXMLNode("longitude", parent=h4)
h6 <- newXMLNode("latitude", parent=h4)
h7 <- newXMLNode("altitude", parent=h4)
h8 <- newXMLNode("tilt", parent=h4)
h9 <- newXMLNode("heading", parent=h4)
h10 <- newXMLNode("altitudeMode", parent=h4)
h11 <- newXMLNode("Folder", "allCtryData", parent=allCtryData.kml[["Document"]])
txtc <- sprintf("<Placemark id='%.0f'><name>%s</name><Point><coordinates>%.5f,%.5f,%.0f</coordinates></Point></Placemark>",
			Temp[,1],allCtryData[[6]],coordinates(allCtryData)[,1], coordinates(allCtryData)[,2], rep(0, length(allCtryData)))
parseXMLAndAdd(txtc, parent = h4)
saveXML(allCtryData.kml, file="C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/allCtryData.kml")
```
# The KML Document does not have the information we are looking for, so we try other methods to generate the correct KML Document

## View the KML Document

```{r}
allCtryData.kml
```

## Save the KML Document as an XML Document so that we can review it

```{r}
saveXML(allCtryData.kml, file="C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/allCtryData.xml")
```

## Add Placemarks to the KML Doc

```{r}
addPlacemark =  function(lat, lon, id, label, parent){
  newXMLNode("Placemark",
 	     newXMLNode("name", label),
	     newXMLNode("Point",
                       newXMLNode("coordinates",
				paste(lon, lat, 0, sep = ","))),
	     attrs = c(id = id), parent = parent)
}
```

## Create Population Scales for the visualization in KML

```{r}
popScales = as.character(1+ c(0.25, 0.75, 1.5, 3, 7))
icon = rep (sprintf("yor%dball.png",
		seq(along = levels( InfMortDiscrete))),
		each = length(levels(InfMortDiscrete)))
scale = rep(popScales, length(levels(popDiscrete)))
```

## Create the Circular Points for display in KML and apply to allCtryData

```{r}
ballStyles = mapply(function(scale, icon)
		list(IconStyle =
		list(scale = scale,
		Icon = c(href = icon))),
		scale, icon, SIMPLIFY=TRUE)
g = expand.grid(seq(along = levels(InfMortDiscrete)),
	seq(along = levels(popDiscrete)))
names(ballStyles) = sprintf("ball_%d-%d", g[,2], g[,1])
ctryStyle = sprintf("#ball_%d-%d", InfMortDiscrete, popDiscrete)
kmlPoints(allCtryData, docStyles = ballStyles, style = ctryStyle)
```

## Create Point Descriptions, the information that displays when you click on a country in Google Earth

```{r}
ptDescriptions =
 sprintf(paste(
"<table><tr><td>Country:</td><td>%s</td></tr>",
"<tr><td>Infant Mortality:</td>",
"<td>%s per 1,000 live births</td></tr>",
"<tr><td>Population:</td><td>%s</td></tr></table>"),
	allCtryData$name, allCtryData$infMort, allCtryData$pop)
```

## Create the final KML Document

```{r}
docName = "Infant Mortality"
docDescription = "2014 CIA Factbook"
folderName = "Countries"
doc = kmlPoints(allCtryData, docName = docName,
	docDescription = docDescription,
	docStyles = ballStyles,
	folderName = folderName,
	style = ctryStyle,
	description = ptDescriptions,
	ids = allCtryData$ctry,
	.names = allCtryData$name)
```

# kmlPoints does not work, so the code in the text will not allow us to create the document and the following code in the text does not work either

## Create the Legend for the Document and save

```{r}
kmlLegend(x = 20, y = 20, title = "Infant Mortality",
	legend = levels(InfMortDiscrete), fill = cols,
	text.col= "white", dims = c(100, 108),
	parent = doc)
saveXML(doc, file = "C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/ctryFancy.kml")
```

## Create KML Document that should display all of the information for Google Earth

```{r}
makeBaseDocument = 
  function(docName = "Infant mortality",
	   docDesc = "2014 CIA Factbook",
	   lat = 43, lon = -121, alt = 4100000,
	   tilt = 0, heading = 0)
{
 doc = newXMLDoc()
 rootNode = newXMLNode("kml", doc = doc)
 DocNode = newXMLNode("Document", parent = rootNode)
 newXMLNode("name", docName , parent = DocNode)
 newXMLNode("description", docDesc, parent = DocNode)
 LANode = newXMLNode("LookAt", parent = DocNode)
 newXMLNode("longitude", lon, parent = LANode)
 newXMLNode("latitude", lat, parent = LANode)
 newXMLNode("tilt", tilt, parent = LANode)
 newXMLNode("heading", heading, parent = LANode)
 newXMLNode("altitudeMode", "absolute", parent = LANode)
 newXMLNode("Folder", parent = DocNode)
 return(doc)
}
baseDoc = makeBaseDocument()
baseDoc
```

## Add Placemarks to the XML Document and save

```{r}
addPlacemark =  function(lat, lon, id, label, parent, style = NULL, desc = NULL){
  newXMLNode("Placemark",
 	     newXMLNode("name", label),
	     newXMLNode("Point",
                       newXMLNode("coordinates",
				paste(lon, lat, 0, sep = ","))),
	     attrs = c(id = id), parent = parent)
}
root = xmlRoot(baseDoc)
folder = root[["Document"]][["Folder"]]
mapply(addPlacemark,
       lat = allCtryData$latitude, lon = allCtryData$longitude,
       id = allCtryData$ctry, label = allCtryData$name,
       parent = folder, style = ctryStyle, desc = ptDescriptions)
makeStyleNode = function(styleInfo, id){
  st = newXMLNode("Style", attrs = c("id" = id))
  newXMLNode("IconStyle", 
            newXMLNode("scale", styleInfo$IconStyle$scale),
            newXMLNode("icon", styleInfo$IconStyle$Icon),
		parent = st)
  return(st)
}
styleNodes = mapply(makeStyleNode, ballStyles, names(ballStyles))
addChildren(root[["Document"]], kids = styleNodes, at = 3)
saveXML(baseDoc, file = "C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/countryMashup.kml")
```

## countryMashup.kml does not produce the desired output in Google Earth either, so we keep trying other methods to get the desired output

## Save the Placemarks as a vectorized string

```{r}
kmlTxt = sprintf("<Placemark><Point><coordinates>%.3f,%.3f,0
		</coordinates></Point></Placemark>",
		allCtryData$longitude, allCtryData$latitude)
```

## Create the XML Document using a different method, should solve the errors we are encountering with the method in the text

## Create KML Document

```{r}
LAST<- newXMLNode("kml")
h2 <- newXMLNode("Document", parent = LAST)
h3 <- newXMLNode("name", "allCtryData", parent = h2)
h4 <- newXMLNode("LookAt", parent=h2)
txtc <- sprintf("<Placemark id='%.0f'><name>%s</name>  <styleUrl>%s</styleUrl>       <Point><coordinates>%.5f,%.5f,%.0f</coordinates></Point></Placemark>",
                Temp[,1],allCtryData[[6]],ctryStyle,coordinates(allCtryData)[,1], coordinates(allCtryData)[,2], rep(0, length(allCtryData)))
parseXMLAndAdd(txtc, parent = h4)
```

## Save the KML Document as an XML Document

```{r}
saveXML(LAST, file="C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/countryData2.kml")
```

## Turn the Pushpin Styles into a list

```{r}
b <-rep(1:5, each=5)
a <- rep(seq(1, 5, 1), 5)
a <- b <- as.factor(a)
ctryStyle1 = sprintf("ball_%d-%d", InfMortDiscrete, popDiscrete)
balls= sprintf("ball_%d-%d", a, b)
icon = rep (sprintf("yor%dball.png",
                    seq(along = levels( InfMortDiscrete))),
            each = length(levels(InfMortDiscrete)))
```

## Create a KML Document that will have the Pushpin styles

```{r}
SS<- newXMLNode("kml")
h2 <- newXMLNode("Document", parent = SS)
h3 <- newXMLNode("name", "allCtryData", parent = h2)
h4 <- newXMLNode("LookAt", parent=h2)
txtc1 <- sprintf("<Style id='%s'> <IconStyle> <scale>%s  </scale>  <Icon>%s </Icon> </IconStyle></Style> ",
                 balls,scale,icon)
txtc <- sprintf("<Placemark id='%.0f'><name>%s</name>  <styleUrl>%s</styleUrl>       <Point><coordinates>%.5f,%.5f,%.0f</coordinates></Point></Placemark>",
                Temp[,1],allCtryData[[6]],ctryStyle,coordinates(allCtryData)[,1], coordinates(allCtryData)[,2], rep(0, length(allCtryData)))
parseXMLAndAdd(txtc, parent = h4)
parseXMLAndAdd(txtc1, parent = h2)
saveXML(SS, file="C:/Users/Robert/Desktop/STAT6306_McGee/Project/FinalProject/countryData3.kml")
```

## Compare two different ways to add the placemarks
# Create the Placemarks using strings

```{r}
addPlacemarks.fast = 
function(lon, lat, parent)
{
 txt = sprintf("<Placemark><Point><coordinates>%.3f,%.3f,0
                </coordinates></Point></Placemark>",
               lon, lat)
 parseXMLAndAdd( paste(txt, collapse = ""),parent)
}
```

# Create the Placemarks using XMLNodes

```{r}
makePM = function(x, y, parent){
   newXMLNode("Placemark",
              newXMLNode("Point",
                          newXMLNode("coordinates",
                                     paste(x, y, 0, sep=","))),
                          parent = parent)
}
addPlacemarks.slow = 
function(lon, lat, parent)
{
 mapply(makePM, x = lon, y = lat, parent =parent)
}
```

# Set up the document and the inputs to the functions

```{r}
doc = newXMLDoc()
root = newXMLNode("kml", doc = doc)
folder = newXMLNode("Folder", parent = root)
lons = rep(allCtryData$longitude, 10)
lats = rep(allCtryData$latitude, 10)
```

# Time the XMLNodes method of creating the Placemarks

```{r}
system.time(invisible(
                    addPlacemarks.slow(lons, lats, folder)))
```

# Time the strings method of creating the Placemarks

```{r}
rm(doc)
doc = newXMLDoc()
root = newXMLNode("kml", doc = doc)
folder = newXMLNode("Folder", parent = root)
system.time(invisible(
            addPlacemarks.fast(lons, lats, folder)))
```

## The strings method is much faster than the XMLNodes method