---
title: "Assignment 6"
author: "RobertFarrow"
date: "November 18, 2015"
output: html_document
---

## Load the packages for this analysis

```{r}
library(caret)
library(rpart)
library(Sleuth3)
library(rpart.plot)
```

## Load the data set and set it to an object

```{r}
Esteem <- ex1223
```

## Look at the data, create the Log(Income2005) variable, and then create the revised data set with the variables we need

```{r}
head(Esteem)
summary(Esteem)
sapply(Esteem, sd)
Esteem$LIncome2005 <- log(Esteem$Income2005)
REsteem <- Esteem[,c("Esteem1", "LIncome2005", "AFQT", "Gender", "Educ")]
REsteem$Esteem1[REsteem$Esteem1 == "3"] <- 0
REsteem$Esteem1[REsteem$Esteem1 == "2"] <- 0
REsteem$Esteem1[REsteem$Esteem1 == "1"] <- 0
REsteem$Esteem1[REsteem$Esteem1 == "4"] <- 1
```

## Create the training and test data sets

```{r}
train_idx <- sample(1:nrow(REsteem),1000,replace=FALSE)
training <- REsteem[train_idx,] # select all these rows
test <- REsteem[-train_idx,] # select all but these rows
```

## Create confusion table

```{r}
train(Esteem1~., data=training, method='pls')
```

## Logistic Regression on the data

```{r}
myLogit <- glm(Esteem1 ~ LIncome2005 + AFQT + Gender + Educ, data=REsteem)
myLogit
summary(myLogit)
```

None of the variables are statistically significant at the alpha = 0.05 level of significance.

## Classification Tree for the data

```{r}
TREsteem <- rpart(myLogit)
TREsteem
plot(TREsteem)
text(TREsteem)
prp(TREsteem)
TREsteem1 <- rpart(myLogit, method='class', cp=0.0001)
printcp(TREsteem)
plotcp(TREsteem)
```

The classification tree does not need to be pruned.  Based off of the cp plot, the classification tree is the best one available.  AFQT greater than or equal to 26.618 is the best classification.